---
title: "Retrospective Longitudinal Analysis of Spectral Features Reveals Divergent Vocal Development Patterns for Treble and Non-Treble Singers"
author: "Reuben Walker"
date: (`r Sys.Date()`)
output: pdf_document
toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_libraries}
suppressPackageStartupMessages({
  suppressWarnings({
    library(ggplot2)
    library(dplyr)
    library(ggpubr)
    library(sjPlot)
    library(lmerTest)
    library(ggeffects)
    library(ggResidpanel)
    library(performance)
    library(emmeans)
    library(knitr)
    library(patchwork)
    library(viridis)
  })
})
```
Research Questions: 

Do classical singers develop similarly over the course of their conservatory training or does their spectral development exhibit differences based on a grouping in female/countertenor or male voice types?

Does breathiness decrease among female students with vocal training?

Statistical Analysis

In each of the nine cases incorporating our three dependent variables and three sung tasks, we performed a linear mixed model with the predictor variables years of study and Voice_Group (male or female/countertenor), with specific emphasis on the interaction between the two predictors. A control variable delta SPL (dSPL) was measured from a reference sample. To account for inter-subject variation in baseline and development, random slopes and intercepts were specified for the individual subjects. We iteratively simplified the models by removing non-significant terms beginning with the interaction term. A linear mixed model was then performed on the female subset for each sung task to investigate increases in CPPs. In order to adjust for multiple comparisons, we adjusted our significance level using the Benjamini-Hochberg correction. We calculated estimated marginal means from the resulting models to examine differences between Voice_Group at different intervals of training.  

# Sustained high phonation
## Load and Prepare Data
```{r}
#Set path
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
#These are the high sustained samples
klang <- read.csv("Klang2_JASA_Final.csv", fileEncoding = "UTF-8")
klang <- klang[c("yearDiff", "H1H2LTAS", "alphaRatio", "CPPs", "Jahr", "Stimmfach", "Voice.Type", "geschlecht", "alter", "id", "dSPL")]
names(klang)[names(klang) == "geschlecht"] <- "Gender"
names(klang)[names(klang) == "yearDiff"] <- "Years"
names(klang)[names(klang) == "alter"] <- "Age"
klang$Voice_Group[klang$Stimmfach == "Sop/Mezzo/Alt"] <- "Treble"
klang$Voice_Group[klang$Stimmfach == "Ten/Bar/Bass"] <- "Non-Treble"
klang$Gender[klang$Gender == "männl."] <- "Male"
klang$Gender[klang$Gender == "weibl."] <- "Female"

```
## Filtering Data
```{r}
# Step 1: Count total unique male and female students
total_counts <- klang %>%
  group_by(Gender) %>%
  summarise(Total_Students = n_distinct(id))

# Step 2: Filter to only include students with Years <= 4
klang_filtered <- klang %>%
  filter(Years <= 4)

# Count unique male and female students after filtering
filtered_counts <- klang_filtered %>%
  group_by(Gender) %>%
  summarise(Filtered_Students = n_distinct(id))

# Step 3: Remove students with only one recording
id_counts <- klang_filtered %>%
  count(id)  # Count occurrences of each student (id)

valid_ids <- id_counts %>%
  filter(n > 1) %>%
  pull(id)  # Get list of students who appear more than once

klang_final <- klang_filtered %>%
  filter(id %in% valid_ids)  # Keep only students with multiple recordings

# Count unique male and female students after final filtering
final_counts <- klang_final %>%
  group_by(Gender) %>%
  summarise(Final_Students = n_distinct(id))

# Print results
print(total_counts)      # Total students by gender
print(filtered_counts)   # After Years <= 4 restriction
print(final_counts)      # After removing single-recording students

# Update the klang dataframe to keep only the final filtered version
klang <- klang_final
```
```{r}
# Split the data into Male/Female and remove Gender column
klang_f <- subset(klang, Gender %in% c("Female"))
# remove Gender
klang_f <- subset(klang_f, select = c("Years", "H1H2LTAS", "alphaRatio", "CPPs", "Jahr", "Gender", "Voice.Type", "Age", "id", "dSPL"))

klang_m <- subset(klang, Gender %in% c("Male"))
klang_m <- subset(klang_m, select = c("Years", "H1H2LTAS", "alphaRatio", "CPPs", "Jahr", "Gender", "Voice.Type", "Age", "id","dSPL"))
```
## H1H2LTAS

```{r}
# Here we take a look at the correlation matrix for our different metrics.
numeric_columns <- klang_f[, c("Years", "H1H2LTAS", "alphaRatio", "CPPs")]
# Omit na values:
cor(na.omit(numeric_columns))
# Repeat for male voices
numeric_columns <- klang_m[, c("Years", "H1H2LTAS", "alphaRatio", "CPPs")]
# Omit na values:
cor(na.omit(numeric_columns))
```

It makes intuitive sense that H1H2LTAS has a strong negative correlation to alpha ratio in women because H2 is above 1000 Hz.
The direction of the Years correlations fits with the hypothesese:
H1H2LTAS tends to increase for women, decrease for men
Alpha Ratio increases for men, seems to remain constant for women
CPPs tends to increase for both men and women

Here we take a look at our linear mixed models:
```{r}
#Let's make treble the reference group
klang$Voice_Group <- factor(klang$Voice_Group)  # convert to factor
klang$Voice_Group <- relevel(klang$Voice_Group, ref = "Treble")  # now relevel
```

```{r}
fit0_H <- lmer(H1H2LTAS~Years*Voice_Group + dSPL + (Years | id), data=klang, na.action="na.omit",REML = T)
anova(fit0_H)
```
We remove dSPL from the model as it was non-significant.
```{r}
fit1_H <- lmer(H1H2LTAS~Years*Voice_Group + (Years | id), data=klang, na.action="na.omit",REML = T)
anova(fit1_H)
```
As the interaction is significant we leave years
```{r}
margin1 <- ggpredict(fit1_H, c("Years", "Voice_Group"), ci_level=0.95)
margin1
```
The directionality of the two groups is clear (female increase, male decrease), though the confidence intervals overlap.
```{r}

emm1_H <- emmeans::emmeans(fit1_H, ~ Years:Voice_Group)
pairs(emm1_H)
p.fit1_H <- ggResidpanel::resid_panel(fit1_H,
              plots = c("resid", "qq", "hist", "index"),
              smoother = TRUE,
              qqbands = TRUE,
              title.opt = TRUE)
p.fit1_H
```

```{r}
summary(fit1_H)
fig_h1 <- plot_model(fit1_H, type = "pred", terms = c("Years", "Voice_Group"), show.data= TRUE,
           colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size=15) +
  labs(x = "Years Studied",
		y = "H1H2LTAS (dB)",
		title = "H1H2LTAS Sustained High Marginal Means") 
fig_h1
ggsave("H1H2LTASMarginal_JASA_T_test_Final.pdf", width=9.25, height=5.71)
```
```{r}
performance::icc(fit1_H)
```
```{r}
#Individual Trajectories

# Clean and prep
klang_clean <- klang[!is.na(klang$H1H2LTAS), ]
klang_clean$fitted <- fitted(fit1_H)

# Ensure factors
klang_clean$id <- as.factor(klang_clean$id)
klang_clean$Voice_Group <- as.factor(klang_clean$Voice_Group)

fig_h1 <- plot_model(fit1_H, type = "pred", terms = c("Years", "Voice_Group"), show.data= FALSE,
           colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size=15) +
  labs(x = "Years Studied",
		y = "H1H2LTAS (dB)",
		title = "H1H2LTAS Sustained High Individual Slopes") 
#fig_h1
#ggsave("H1H2LTASMarginal_JASA_T_test_Final.pdf", width=9.25, height=5.71)

# Plot all subjects, colored by Voice_Group
fig_slopes1 <- fig_h1 + geom_line(data = klang_clean, 
            aes(x = Years, y = fitted, group = id, color = Voice_Group), 
            alpha = 0.3, linewidth = 0.7, inherit.aes = FALSE) #+

ggsave("H1H2LTASMarginal_JASA_Individual_Final.pdf", width=9.25, height=5.71)

```
```{r}
library(lme4)
library(broom.mixed)

# Extract random slopes
re_slopes <- ranef(fit1_H)$id
re_df <- data.frame(id = rownames(re_slopes), slope = re_slopes$Years)

# Join Voice_Group info
re_df <- left_join(re_df, unique(klang_clean[, c("id", "Voice_Group")]), by = "id")

# Plot random slopes by Voice Group
ggplot(re_df, aes(x = reorder(id, slope), y = slope, fill = Voice_Group)) +
  geom_col(show.legend = TRUE) +
  coord_flip() +
  scale_fill_viridis_d(option = "C") +
  labs(title = "Random Slopes by Subject",
       x = "Subject ID", y = "Deviation in Years Slope") +
  theme_minimal(base_size = 14)
```
## Alpha-Ratio

```{r}
#Let's make non-treble the reference group
#klang$Voice_Group <- factor(klang$Voice_Group)  # convert to factor
klang$Voice_Group <- relevel(klang$Voice_Group, ref = "Non-Treble")  # now relevel
```

```{r}
#Same process for Alpha Ratio. 
fit0_a <- lmer(alphaRatio~Years*Voice_Group +dSPL + (Years | id), data=klang, na.action="na.omit",REML=T)
anova(fit0_a)

```
Simplify removing the insignificant dSPL term
```{r}
#Same process for Alpha Ratio. 
fit1_a <- lmer(alphaRatio~Years*Voice_Group + (Years | id), data=klang, na.action="na.omit",REML=T)
anova(fit1_a)

```
Again the interaction term is significant, so we don't reduce further.
```{r}
margin1 <- ggpredict(fit1_a, c("Years", "Voice_Group"), ci_level=0.95)
margin1
```
For alpha ratio, the male means in year four have increased beyond the confidence interval at the beginning of studies.
The female means have not changed.
```{r}

emm1_a <- emmeans::emmeans(fit1_a, ~ Years:Voice_Group)
pairs(emm1_a)
p.fit1_a <- ggResidpanel::resid_panel(fit1_a,
              plots = c("resid", "qq", "hist", "index"),
              smoother = TRUE,
              qqbands = TRUE,
              title.opt = TRUE)
p.fit1_a
```

```{r}
summary(fit1_a)
fig_a1 <- plot_model(fit1_a, type = "pred", terms = c("Years", "Voice_Group"), show.data= TRUE,
           colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size=15) +
  labs(x = "Years Studied",
		y = "Alpha Ratio (dB)",
		title = "Alpha Ratio Sustained High Marginal Means")
fig_a1
ggsave("AlphaMarginal_JASA_T_test_Final.pdf", width = 6, height = 4, dpi = 300)

```
```{r}
performance::icc(fit1_a)
```

```{r}
# INDIVIDUAL SLOPES clean and prep
klang_clean <- klang[!is.na(klang$alphaRatio), ]
klang_clean$fitted <- fitted(fit1_a)

# Ensure factors
klang_clean$id <- as.factor(klang_clean$id)
klang_clean$Voice_Group <- as.factor(klang_clean$Voice_Group)

fig_a1 <- plot_model(fit1_a, type = "pred", terms = c("Years", "Voice_Group"), show.data= FALSE,
           colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size=15) +
  labs(x = "Years Studied",
		y = "Alpha Ratio (dB)",
		title = "Alpha Ratio Sustained High Individual Slopes") 
#fig_h1
#ggsave("H1H2LTASMarginal_JASA_T_test_Final.pdf", width=9.25, height=5.71)

# Plot all subjects, colored by Voice_Group
fig_slopes2 <- fig_a1 + geom_line(data = klang_clean, 
            aes(x = Years, y = fitted, group = id, color = Voice_Group), 
            alpha = 0.3, linewidth = 0.7, inherit.aes = FALSE) #+

ggsave("AlphaMarginal_JASA_Individual_Final.pdf", width=9.25, height=5.71)

```
## CPPS

```{r}
#Let's make treble the reference group
klang$Voice_Group <- relevel(klang$Voice_Group, ref = "Treble")  # now relevel
```

```{r}
#Same process for CPPs. _Review interaction term
fit0_c <- lmer(CPPs~Years*Voice_Group + dSPL + (Years | id), data=klang, na.action="na.omit",REML=T)
anova(fit0_c)
```
```{r}
#We can remove the interaction term
fit1_c <- lmer(CPPs~Years + Voice_Group + dSPL + (Years | id), data=klang, na.action="na.omit",REML=T)
anova(fit1_c)
```
Now we can remove dSPL 
```{r}

fit2_c <- lmer(CPPs~Years + Voice_Group + (Years | id), data=klang, na.action="na.omit",REML=T)
anova(fit2_c)

```

We have significant effects for Years and Voice_Group
```{r}
margin1 <- ggpredict(fit2_c, c("Years", "Voice_Group"), ci_level=0.95)
margin1

emm1_c <- emmeans::emmeans(fit2_c, ~ Years:Voice_Group)
pairs(emm1_c)
p.fit2_c <- ggResidpanel::resid_panel(fit2_c,
              plots = c("resid", "qq", "hist", "index"),
              smoother = TRUE,
              qqbands = TRUE,
              title.opt = TRUE)
p.fit2_c
```

```{r}
summary(fit2_c)
fig_c1 <- plot_model(fit2_c, type = "pred", terms = c("Years", "Voice_Group"), show.data= TRUE,
           colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size=15) +
  labs(x = "Years Studied",
		y = "CPPs",
		title = "CPPs Sustained High Marginal Means") +
  theme(aspect.ratio = 1)
fig_c1
ggsave("CPPsMarginal_JASA_T_test_Final.pdf", width=9.25, height=5.71)
```
```{r}
performance::icc(fit2_c)
```

Secondary research question: Do female voices became less breathy?
```{r}
fit0_c_f <- lmer(CPPs~Years + dSPL + (Years | id), data=klang_f, na.action="na.omit",REML=T)
anova(fit0_c_f)
```
dSPL can be removed
```{r}
fit1_c_f <- lmer(CPPs~Years + (Years | id), data=klang_f, na.action="na.omit",REML=T)
summary(fit1_c_f)
```

```{r}
# INDIVIDUAL SLOPES clean and prep
klang_clean <- klang[!is.na(klang$CPPs), ]
klang_clean$fitted <- fitted(fit2_c)

# Ensure factors
klang_clean$id <- as.factor(klang_clean$id)
klang_clean$Voice_Group <- as.factor(klang_clean$Voice_Group)

fig_c1 <- plot_model(fit2_c, type = "pred", terms = c("Years", "Voice_Group"), show.data= FALSE,
           colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size=15) +
  labs(x = "Years Studied",
		y = "CPPs (dB)",
		title = "CPPs Sustained High Individual Slopes") 
#fig_h1
#ggsave("H1H2LTASMarginal_JASA_T_test_Final.pdf", width=9.25, height=5.71)

# Plot all subjects, colored by Voice_Group
fig_slopes3 <- fig_c1 + geom_line(data = klang_clean, 
            aes(x = Years, y = fitted, group = id, color = Voice_Group), 
            alpha = 0.3, linewidth = 0.7, inherit.aes = FALSE) #+

ggsave("CPPsMarginal_JASA_Individual_Final.pdf", width=9.25, height=5.71)
```

Benjamini-Hochberg Correction
We take the p-values from all of our initial linear mixed models:
```{r}
p_values_H <- summary(fit0_H)$coefficients[, "Pr(>|t|)"]
p_values_a <- summary(fit0_a)$coefficients[, "Pr(>|t|)"]
p_values_c <- summary(fit0_c)$coefficients[, "Pr(>|t|)"]
p_values_c_f <- summary(fit0_c_f)$coefficients[, "Pr(>|t|)"]

# Organize p-values into a data frame
p_values_df <- data.frame(
  H1H2LTAS = p_values_H,
  AlphaRatio = p_values_a,
  CPPs = p_values_c#,
  #Model4 = p_values_c_f
)

# Combine all p-values into a single vector
all_p_values <- c(p_values_H, p_values_a, p_values_c)

#Save the unadjusted for final comparison:
p_2 <- all_p_values
p_2_f <- p_values_c_f
```

# Repertoire: Avezzo a vivere

Now for the repertoire sample "Avezzo a vivere"

## Load and Prepare Data
```{r}
klang6 <- read.csv("Klang6_JASA_Final.csv", fileEncoding = "UTF-8")
klang6 <- klang6 %>%
  select(yearDiff, H1H2LTAS, alphaRatio, CPPs, Jahr, Stimmfach, Voice.Type, geschlecht, alter, id, dSPL) %>%
  rename(Years = yearDiff, Gender = geschlecht, Age = alter)

# Create Voice Grouping
klang6 <- klang6 %>%
  mutate(
    Voice_Group = case_when(
      Stimmfach == "Sop/Mezzo/Alt" ~ "Treble",
      Stimmfach == "Ten/Bar/Bass" ~ "Non-Treble"
    ),
    Gender = recode(Gender, "männl." = "Male", "weibl." = "Female")
  )
```

## Filtering Data
```{r}
# Step 1: Count total unique students by gender
total_counts <- klang6 %>% group_by(Gender) %>% summarise(Total_Students = n_distinct(id))

# Step 2: Filter students with Years <= 4
klang6_filtered <- klang6 %>% filter(Years <= 4)
filtered_counts <- klang6_filtered %>% group_by(Gender) %>% summarise(Filtered_Students = n_distinct(id))

# Step 3: Remove students with only one recording
valid_ids <- klang6_filtered %>% count(id) %>% filter(n > 1) %>% pull(id)
klang6_final <- klang6_filtered %>% filter(id %in% valid_ids)
final_counts <- klang6_final %>% group_by(Gender) %>% summarise(Final_Students = n_distinct(id))

# Print Results
print(total_counts)
print(filtered_counts)
print(final_counts)

# Update dataset
klang6 <- klang6_final
```

## H1H2LTAS Analysis

```{r}
#Let's make treble the reference group
klang6$Voice_Group <- factor(klang6$Voice_Group)  # convert to factor
klang6$Voice_Group <- relevel(klang6$Voice_Group, ref = "Treble")  # now relevel
```

```{r}
fit0_H <- lmer(H1H2LTAS ~ Years * Voice_Group + dSPL + (Years | id), data = klang6, na.action = "na.omit", REML = TRUE)
anova(fit0_H)
```
Remove the dSPL term:
```{r}
fit1_H <- lmer(H1H2LTAS ~ Years * Voice_Group + (Years | id), data = klang6, na.action = "na.omit", REML = TRUE)
anova(fit1_H)
# Marginal Means
margin1 <- ggpredict(fit1_H, c("Years", "Voice_Group"), ci_level = 0.95)
print(margin1)

# Pairwise Comparisons
emm0_H <- emmeans(fit1_H, ~ Years:Voice_Group)
pairs(emm0_H)

# Residual Diagnostics
p.fit0_H <- ggResidpanel::resid_panel(fit1_H, plots = c("resid", "qq", "hist", "index"), smoother = TRUE, qqbands = TRUE)
print(p.fit0_H)

# Plot Model Predictions
fig_h2 <- plot_model(fit1_H, type = "pred", terms = c("Years", "Voice_Group"), show.data = TRUE,
                      colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size = 15) +
  labs(x = "Years Studied", y = "H1H2LTAS (dB)", title = "H1H2LTAS Repertoire Marginal Means")
fig_h2
ggsave("H1H2LTASAvezzoMarginal_JASA_T_test_Final.pdf", width = 9.25, height = 5.71)
```
```{r}
summary(fit1_H)
performance::icc(fit1_H)
```
```{r}
# INDIVIDUAL SLOPES clean and prep
klang_clean <- klang6[!is.na(klang6$H1H2LTAS), ]
klang_clean$fitted <- fitted(fit1_H)

# Ensure factors
klang_clean$id <- as.factor(klang_clean$id)
klang_clean$Voice_Group <- as.factor(klang_clean$Voice_Group)

fig_h1 <- plot_model(fit1_H, type = "pred", terms = c("Years", "Voice_Group"), show.data= FALSE,
           colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size=15) +
  labs(x = "Years Studied",
		y = "H1H2LTAS (dB)",
		title = "H1H2LTAS Repertoire Individual Slopes") 
#fig_h1
#ggsave("H1H2LTASMarginal_JASA_T_test_Final.pdf", width=9.25, height=5.71)

# Plot all subjects, colored by Voice_Group
fig_h1 + geom_line(data = klang_clean, 
            aes(x = Years, y = fitted, group = id, color = Voice_Group), 
            alpha = 0.3, linewidth = 0.7, inherit.aes = FALSE) #+

ggsave("H1H2LTASMarginalRepertoire_JASA_Individual_Final.pdf", width=9.25, height=5.71)
```

## Alpha Ratio Analysis

```{r}
#Let's make treble the reference group
klang6$Voice_Group <- relevel(klang6$Voice_Group, ref = "Non-Treble")  # now relevel
```

```{r}
fit0_a <- lmer(alphaRatio ~ Years * Voice_Group + dSPL + (Years | id), data = klang6, na.action = "na.omit", REML = TRUE)
anova(fit0_a)
```
```{r}
#Remove dSPL
fit1_a <- lmer(alphaRatio ~ Years * Voice_Group + (Years | id), data = klang6, na.action = "na.omit", REML = TRUE)
anova(fit1_a)

# Model Simplification
fit2_a <- update(fit1_a, . ~ . - Years:Voice_Group)
anova(fit2_a)

fit3_a <- update(fit2_a, . ~ . - Years)
anova(fit3_a)

# Compare Models
anova(fit0_a, fit1_a, fit2_a, fit3_a)

# Residual Diagnostics
p.fit3_a <- ggResidpanel::resid_panel(fit3_a, plots = c("resid", "qq", "hist", "index"), smoother = TRUE, qqbands = TRUE)
print(p.fit3_a)

# Plot Model Predictions
plot_model(fit3_a, type = "pred", terms = c("Years", "Voice_Group"), show.data = TRUE,
           colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size = 15) +
  labs(x = "Years Studied", y = "Alpha Ratio (dB)", title = "Alpha Ratio Repertoire Marginal Means")
ggsave("AlphaAvezzoMarginal_JASA_T_test_Final.pdf", width = 9.25, height = 5.71)
```
```{r}
summary(fit3_a)
performance::icc(fit3_a)
```

## CPPs Analysis

```{r}
#Let's make treble the reference group
klang6$Voice_Group <- relevel(klang6$Voice_Group, ref = "Treble")  # now relevel
```

```{r}
fit0_c <- lmer(CPPs ~ Years * Voice_Group + dSPL + (Years | id), data = klang6, na.action = "na.omit", REML = TRUE)
anova(fit0_c)
```
```{r}
#Remove dSPL
fit1_c <- lmer(CPPs ~ Years * Voice_Group + (Years | id), data = klang6, na.action = "na.omit", REML = TRUE)
anova(fit1_c)

# Marginal Means
margin1 <- ggpredict(fit1_c, c("Years", "Voice_Group"), ci_level = 0.95)
print(margin1)

# Pairwise Comparisons
emm1_c <- emmeans(fit1_c, ~ Years:Voice_Group)
pairs(emm1_c)

# Residual Diagnostics
p.fit1_c <- ggResidpanel::resid_panel(fit1_c, plots = c("resid", "qq", "hist", "index"), smoother = TRUE, qqbands = TRUE)
print(p.fit1_c)

# Plot Model Predictions
fig_c2 <- plot_model(fit1_c, type = "pred", terms = c("Years", "Voice_Group"), show.data = TRUE,
                      colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size = 15) +
  labs(x = "Years Studied", y = "CPPs", title = "CPPs Repertoire Marginal Means")
fig_c2
ggsave("CPPsAvezzoMarginal_JASA_T_test_Final.pdf", width = 9.25, height = 5.71)
```
```{r}
summary(fit1_c)
performance::icc(fit1_c)
```

```{r}
fit0_c_f <- lmer(CPPs ~ Years  + dSPL + (Years | id), data = klang6 %>% filter(Gender == "Female"), REML = TRUE)
anova(fit0_c_f)
```
```{r}
#Remove dSPL
fit1_c_f <- lmer(CPPs ~ Years + (Years | id), data = klang6 %>% filter(Gender == "Female"), REML = TRUE)
summary(fit1_c_f)
```

```{r}
p_values_H <- summary(fit0_H)$coefficients[, "Pr(>|t|)"]
p_values_a <- summary(fit0_a)$coefficients[, "Pr(>|t|)"]
p_values_c <- summary(fit0_c)$coefficients[, "Pr(>|t|)"]
p_values_c_f <- summary(fit0_c_f)$coefficients[, "Pr(>|t|)"]

all_p_values <- c(p_values_H, p_values_a, p_values_c)#, p_values_c_f)
p_6 <- all_p_values
p_6_f <- p_values_c_f
```


# Medium Sustained /a/
## Load and Prepare Data
```{r}
# Load data
klang1 <- read.csv("Klang1_JASA_Final.csv", fileEncoding = "UTF-8")
klang1 <- klang1[c("yearDiff", "H1H2LTAS", "alphaRatio", "CPPs", "Jahr", "Stimmfach", "Voice.Type", "geschlecht", "alter", "id", "dSPL")]
names(klang1)[names(klang1) == "geschlecht"] <- "Gender"
names(klang1)[names(klang1) == "yearDiff"] <- "Years"
names(klang1)[names(klang1) == "alter"] <- "Age"
klang1$Voice_Group[klang1$Stimmfach == "Sop/Mezzo/Alt"] <- "Treble"
klang1$Voice_Group[klang1$Stimmfach == "Ten/Bar/Bass"] <- "Non-Treble"
klang1$Gender[klang1$Gender == "männl."] <- "Male"
klang1$Gender[klang1$Gender == "weibl."] <- "Female"

```
## Filtering Data
```{r}
# Step 1: Count total unique male and female students
total_counts <- klang1 %>%
  group_by(Gender) %>%
  summarise(Total_Students = n_distinct(id))

# Step 2: Filter to only include students with Years <= 4
klang1_filtered <- klang1 %>%
  filter(Years <= 4)

# Count unique male and female students after filtering
filtered_counts <- klang1_filtered %>%
  group_by(Gender) %>%
  summarise(Filtered_Students = n_distinct(id))

# Step 3: Remove students with only one recording
id_counts <- klang1_filtered %>%
  count(id)  # Count occurrences of each student (id)

valid_ids <- id_counts %>%
  filter(n > 1) %>%
  pull(id)  # Get list of students who appear more than once

klang1_final <- klang1_filtered %>%
  filter(id %in% valid_ids)  # Keep only students with multiple recordings

# Count unique male and female students after final filtering
final_counts <- klang1_final %>%
  group_by(Gender) %>%
  summarise(Final_Students = n_distinct(id))

# Print results
print(total_counts)      # Total students by gender
print(filtered_counts)   # After Years <= 4 restriction
print(final_counts)      # After removing single-recording students

# Update the klang1 dataframe to keep only the final filtered version
klang1 <- klang1_final
```
## H1H2LTAS
```{r}
#Let's make treble the reference group
klang1$Voice_Group <- factor(klang1$Voice_Group)  # convert to factor
klang1$Voice_Group <- relevel(klang1$Voice_Group, ref = "Treble")  # now relevel
```

```{r}
fit0_H <- lmer(H1H2LTAS ~ Years * Voice_Group + dSPL + (Years | id), data = klang1, REML = TRUE)
anova(fit0_H)
```
```{r}
#After significance correction, not significant: Remove dSPL
fit1_H <- lmer(H1H2LTAS ~ Years * Voice_Group + (Years | id), data = klang1, REML = TRUE)
anova(fit1_H)

#After significance correction, not significant: Remove interaction
fit2_H <- update(fit1_H, . ~ . - Years:Voice_Group)
anova(fit2_H)
#Remove years
fit3_H <- update(fit2_H, . ~ . - Years)
anova(fit3_H)

anova(fit0_H, fit1_H, fit2_H, fit3_H, digits = 10, test = "LRT")

p.fit3_H <- resid_panel(fit3_H, plots = c("resid", "qq", "hist", "index"), smoother = TRUE, qqbands = TRUE, title.opt = TRUE)
p.fit3_H
```

```{r}
plot_model(fit3_H, type = "pred", terms = c("Years", "Voice_Group"), show.data = TRUE, colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size = 15) +
  labs(x = "Years Studied", y = "H1H2LTAS (dB)", title = "H1H2LTAS Sustained Medium Marginal Means")
ggsave("H1H2LTASMediumMarginal_JASA_T_test_Final.pdf", width = 9.25, height = 5.71)
```

```{r}
summary(fit3_H)
performance::icc(fit3_H)
```

## Alpha-Ratio
```{r}
#Let's make treble the reference group
klang1$Voice_Group <- relevel(klang1$Voice_Group, ref = "Non-Treble")  # now relevel
```

```{r}
fit0_a <- lmer(alphaRatio ~ Years * Voice_Group + dSPL + (Years | id), data = klang1, REML = TRUE)
anova(fit0_a)
```
```{r}
#Remove dSPL
fit1_a <- lmer(alphaRatio ~ Years * Voice_Group + (Years | id), data = klang1, REML = TRUE)
anova(fit1_a)

fit2_a <- update(fit1_a, . ~ . - Years:Voice_Group)
anova(fit2_a)

fit3_a <- update(fit2_a, . ~ . - Years)
anova(fit3_a)

anova(fit0_a, fit1_a, fit2_a, fit3_a, digits = 10)

p.fit3_a <- resid_panel(fit3_a, plots = c("resid", "qq", "hist", "index"), smoother = TRUE, qqbands = TRUE, title.opt = TRUE)
p.fit3_a
```

```{r}
plot_model(fit3_a, type = "pred", terms = c("Years", "Voice_Group"), show.data = TRUE, colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size = 15) +
  labs(x = "Years Studied", y = "Alpha Ratio (dB)", title = "Alpha Ratio Medium Sustained Marginal Means")
ggsave("AlphaMediumMarginal_JASA_T_test_Final.pdf", width = 9.25, height = 5.71)
```

```{r}
summary(fit3_a)
performance::icc(fit3_a)
```

## CPPs
```{r}
#Let's make treble the reference group
klang1$Voice_Group <- relevel(klang1$Voice_Group, ref = "Treble")  # now relevel
```

```{r}
fit0_c <- lmer(CPPs ~ Years * Voice_Group + dSPL + (Years | id), data = klang1, REML = TRUE)
anova(fit0_c)
```
```{r}
#Remove dSPL
fit1_c <- lmer(CPPs ~ Years * Voice_Group + (Years | id), data = klang1, REML = TRUE)
anova(fit1_c)

margin1 <- ggpredict(fit1_c, c("Years", "Voice_Group"), ci_level = 0.95)
print(margin1)

emm0_c <- emmeans(fit1_c, ~ Years:Voice_Group)
pairs(emm0_c)

p.fit1_c <- resid_panel(fit1_c, plots = c("resid", "qq", "hist", "index"), smoother = TRUE, qqbands = TRUE, title.opt = TRUE)
p.fit1_c
```

```{r}
fig_c3 <- plot_model(fit1_c, type = "pred", terms = c("Years", "Voice_Group"), show.data = TRUE, colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size = 15) +
  labs(x = "Years Studied", y = "CPPs", title = "CPPs Medium Sustained Marginal Means") +
  theme(aspect.ratio = 1)
fig_c3
ggsave("CPPsMediumMarginal_JASA_T_test_Final.pdf", width = 9.25, height = 5.71)
```

```{r}
summary(fit1_c)
performance::icc(fit1_c)
```

```{r}
# INDIVIDUAL SLOPES clean and prep
klang_clean <- klang1[!is.na(klang1$H1H2LTAS), ]
klang_clean$fitted <- fitted(fit1_c)

# Ensure factors
klang_clean$id <- as.factor(klang_clean$id)
klang_clean$Voice_Group <- as.factor(klang_clean$Voice_Group)

fig_h1 <- plot_model(fit1_c, type = "pred", terms = c("Years", "Voice_Group"), show.data= FALSE,
           colors = palette.colors(palette = "Okabe-Ito")) +
  theme_gray(base_size=15) +
  labs(x = "Years Studied",
		y = "CPPs (dB)",
		title = "CPPs Medium Sustained Individual Slopes") 
#fig_h1
#ggsave("H1H2LTASMarginal_JASA_T_test_Final.pdf", width=9.25, height=5.71)

# Plot all subjects, colored by Voice_Group
fig_h1 + geom_line(data = klang_clean, 
            aes(x = Years, y = fitted, group = id, color = Voice_Group), 
            alpha = 0.3, linewidth = 0.7, inherit.aes = FALSE) #+

ggsave("CPPsMarginalMedium_JASA_Individual_Final.pdf", width=9.25, height=5.71)
```



```{r}
fit0_c_f <- lmer(CPPs ~ Years + dSPL + (Years | id), data = klang1 %>% filter(Gender == "Female"), REML = TRUE)
anova(fit0_c_f)
```
```{r}
#Remove dSPL
fit1_c_f <- lmer(CPPs ~ Years + (Years | id), data = klang1 %>% filter(Gender == "Female"), REML = TRUE)
summary(fit1_c_f)
```
```{r}
p_values_H <- summary(fit0_H)$coefficients[, "Pr(>|t|)"]
p_values_a <- summary(fit0_a)$coefficients[, "Pr(>|t|)"]
p_values_c <- summary(fit0_c)$coefficients[, "Pr(>|t|)"]

all_p_values <- c(p_values_H, p_values_a, p_values_c)
p_1 <- all_p_values
p_1_f <- summary(fit0_c_f)$coefficients[, "Pr(>|t|)"]

```



```{r}
# Stack plots with superscripts and a caption
combined_plot <- (fig_c1 + labs(tag = "A")) /
                 (fig_c3 + labs(tag = "B")) /
                 (fig_c2 + labs(tag = "C")) +
  plot_annotation(
    tag_levels = "A"
    #title = "Figure 1"#,
    #caption = "CPPs marginal means for sustained high, repertoire, and medium sustained contexts."
  ) +
  plot_layout(heights = c(1, 1, 1))  # Equal heights

# Save the combined plot with fixed dimensions
#combined_plot
ggsave("combined_CPPs_JASA_test_Final.pdf", plot = combined_plot, width = 6, height = 12, dpi = 300)

```

```{r}
# Stack plots with superscripts and a caption
combined_plot <- (fig_h1 + labs(tag = "A")) /
                 (fig_h2 + labs(tag = "B")) +
  plot_annotation(
    tag_levels = "A"
    #title = "Figure 1"#,
    #caption = "CPPs marginal means for sustained high, repertoire, and medium sustained contexts."
  ) +
  plot_layout(heights = c(1, 1))  # Equal heights

# Save the combined plot with fixed dimensions
#combined_plot
ggsave("combined_H1H2LTAS_JASA_test_Final.pdf", plot = combined_plot, width = 6, height = 8, dpi = 300)
```
```{r}
# Stack plots with superscripts and a caption
combined_plot <- (fig_slopes1 + labs(tag = "A")) /
                 (fig_slopes2 + labs(tag = "B")) /
                 (fig_slopes3 + labs(tag = "C")) +
  plot_annotation(
    tag_levels = "A"
    #title = "Figure 1"#,
    #caption = "CPPs marginal means for sustained high, repertoire, and medium sustained contexts."
  ) +
  plot_layout(heights = c(1, 1, 1))  # Equal heights

# Save the combined plot with fixed dimensions
#combined_plot
ggsave("combined_Slopes.pdf", plot = combined_plot, width = 6, height = 8, dpi = 300)
```

# P-Value Adjustments
```{r}

total_p_values <- c(p_2, p_6, p_1, p_2_f, p_6_f, p_1_f)

# Apply the Benjamini-Hochberg correction using p.adjust()
adjusted_p_values_total <- p.adjust(total_p_values, method = "BH")

#Final DFs
p_values_BH_2 <- data.frame(
  H1H2LTAS = adjusted_p_values_total[1:5],
  AlphaRatio = adjusted_p_values_total[6:10],
  CPPs = adjusted_p_values_total[11:15]#,
)

p_values_BH_6 <- data.frame(
  H1H2LTAS = adjusted_p_values_total[16:20],
  AlphaRatio = adjusted_p_values_total[21:25],
  CPPs = adjusted_p_values_total[26:30]#,
)

p_values_BH_1 <- data.frame(
  H1H2LTAS = adjusted_p_values_total[31:35],
  AlphaRatio = adjusted_p_values_total[36:40],
  CPPs = adjusted_p_values_total[41:45]#,
)


p_values_BH_f <- data.frame(
  Intercept = adjusted_p_values_total[c(46, 49, 52)],
  Years = adjusted_p_values_total[c(47, 50, 53)],
  dSPL = adjusted_p_values_total[c(48, 51, 54)]
)
rownames(p_values_BH_f) <- c("High", "Rep", "Med") 

# Display results with formatted tables
cat("### High Frequency Sustained Phonation\n")
kable(p_values_BH_2, digits = 3)

cat("\n### Medium Frequency Sustained Phonation\n")
kable(p_values_BH_1, digits = 3)

cat("\n### Repertoire\n")
kable(p_values_BH_6, digits = 3)

cat("\n### Female CPPs\n")
kable(p_values_BH_f, digits = 3)

```

